### 2021-09-23 04--blast_classification.Rmd

---
title: "Alternative taxonomic assignment using BLAST"
bibliography: '`r sharedbib::bib_path()`'
output:
  html_document:
    css: style.css
---

```{r setup, include=FALSE}
source('style.R')
```

## Prepare

### Packages used

```{r message=FALSE}
library(dplyr)
library(purrr)
library(furrr)
library(tidyr)
library(readr)
library(ggplot2)
library(sessioninfo)
library(metacoder)
library(vegan)
library(viridis)
library(DT)
library(stringr)
library(qsubmitter)
library(taxize)
```

### Parameters

```{r}
cgrb <- remote_server$new(server = "shell.cgrb.oregonstate.edu", user = "fosterz", port  = 732)
files <- remote_server$new(server = "files.cgrb.oregonstate.edu", user = "fosterz", port  = 732)
remote_repository_path <- "/dfs/Grunwald_Lab/home/fosterz/repositories/rps10_barcode"
```

### Parameters

```{r}
seed <- 1
min_evalue <- '1e-3'
set.seed(seed)
```

## BLAST

A blast-based classification will be useful for the the non-target sequences detection.
It will also be useful for verifying the `dada2`-based classification.

### Make query file

I will make FASTA files of ASV and OTU sequences with header containing their row indexes in the abundance matrix.

```{r}
# ASVs
abundance_asv <- read_csv(file.path('intermediate_data', 'abundance_asv.csv'))
query_seq_path_asv <- file.path('intermediate_data', 'blast_query_asv.fa')
paste0('>asv_', 1:nrow(abundance_asv), '\n', abundance_asv$sequence) %>%
  write_lines(file = query_seq_path_asv)

# OTUs
abundance_otu <- read_csv(file.path('intermediate_data', 'abundance_otu.csv'))
query_seq_path_otu <- file.path('intermediate_data', 'blast_query_otu.fa')
paste0('>otu_', 1:nrow(abundance_otu), '\n', abundance_otu$sequence) %>%
  write_lines(file = query_seq_path_otu)
```

### Run BLAST

First I will need to transfer the file from this computer to the CGRB cluster.

```{r blast_upload_query}
# ASVs
remote_query_path_asv <- file.path(remote_repository_path, 'blast_query_asv.fa')
rsync_push(local_path = query_seq_path_asv, remote_path = remote_query_path_asv, remote = files)

# OTUs
remote_query_path_otu <- file.path(remote_repository_path, 'blast_query_otu.fa')
rsync_push(local_path = query_seq_path_otu, remote_path = remote_query_path_otu, remote = files)
```

Then I can run blast remotely.

```{r blast_run}
out_cols <- 'qseqid sallseqid pident length mismatch gapopen qstart qend sstart send evalue bitscore staxids sskingdoms score sscinames'
run_blast_remotely <- function(query_path, out_path, ...) {
  blast_command <- paste('blastn',
                         "-query", query_path,
                         "-db nt",
                         "-dust no",
                         paste("-evalue", min_evalue),
                         # "-perc_identity 70",
                         paste0("-outfmt '10 ", out_cols, "'"),
                         "-max_hsps 1",
                         "-max_target_seqs 100",
                         '-num_threads 8',
                         '-out', out_path)
  qsub(command = blast_command,
       remote = cgrb,
       remote_cwd = remote_repository_path,
       cores = 8,
       queue = 'bpp@!(uncia)', # For some reason, I was getting errors on uncia, so this avoids that node
       ...)
}

# ASVs
remote_blast_out_asv <- file.path(remote_repository_path, 'blast_result_asv.csv')
run_blast_remotely(query_path = remote_query_path_asv, out_path = remote_blast_out_asv)

# OTUs
remote_blast_out_otu <- file.path(remote_repository_path, 'blast_result_otu.csv')
run_blast_remotely(query_path = remote_query_path_otu, out_path = remote_blast_out_otu)
```

and download the results:

```{r blast_download}
# ASVs
blast_result_path_asv <- file.path('intermediate_data', 'blast_results_asv.csv')
rsync_pull(local_path = blast_result_path_asv, remote_path = remote_blast_out_asv, remote = files)

# OTUs
blast_result_path_otu <- file.path('intermediate_data', 'blast_results_otu.csv')
rsync_pull(local_path = blast_result_path_otu, remote_path = remote_blast_out_otu, remote = files)
```

and read them into R:

```{r}
blast_results_asv <- read_csv(blast_result_path_asv, col_names = strsplit(out_cols, split = ' ')[[1]], col_types = 'ccddddddddddcccc')
blast_results_otu <- read_csv(blast_result_path_otu, col_names = strsplit(out_cols, split = ' ')[[1]], col_types = 'ccddddddddddcccc')
```

There might be some warnings about parsing errors caused by `,` in the `sscinames` column, but these should not affect the analysis since that information is not used and there are no columns after it to mess up.
I will select the best hit for each ASV based on e-value and percent identity:

```{r}
select_best_blast_hit <- function(blast_results) {
  blast_results %>%
    group_by(qseqid) %>%
    filter(evalue == min(evalue)) %>%
    filter(pident == max(pident)) %>%
    filter(row_number() == 1) # break ties by picking first value
}

blast_results_asv <- select_best_blast_hit(blast_results_asv)
blast_results_otu <- select_best_blast_hit(blast_results_otu)
```

Then I can look up the taxonomic info from the NCBI taxonomy database using the taxon ID that blast returns:

```{r}
lookup_tax <- function(blast_results) {
  blast_results$staxids <- sub(blast_results$staxids, pattern = ';.+$', replacement = '')
  classification(as.uid(unique(blast_results$staxids), check = FALSE), db = 'ncbi')
}

blast_class_asv <- lookup_tax(blast_results_asv)
blast_class_otu <- lookup_tax(blast_results_otu)
```

and add that to the results table

```{r}
get_classification <- function(x) {
  if (is.logical(x)) {
    return(NA)
  } else {
    return(paste(x$name, collapse = ';'))
  }
}

blast_results_asv$blast_tax <- map_chr(blast_class_asv, get_classification)[blast_results_asv$staxids]
blast_results_otu$blast_tax <- map_chr(blast_class_otu, get_classification)[blast_results_otu$staxids]
```

and combine that with the abundance matrix

```{r}
add_to_abund <- function(abundance, blast_results) {
  blast_results %>%
    ungroup() %>%
    transmute(sequence = abundance$sequence[as.numeric(sub(qseqid, pattern = '^asv_|otu_', replacement = ''))],
              blast_pid = pident,
              blast_cov = c(qend - qstart) / nchar(sequence) * 100,
              blast_tax = blast_tax) %>%
    right_join(abundance, by = 'sequence')
}

abundance_asv <- add_to_abund(abundance_asv, blast_results_asv)
abundance_otu <- add_to_abund(abundance_otu, blast_results_otu)
```

Finally, lets save that modified matrix for further analyses.
Note that this overwrites the abundance matrix.

```{r}
write_csv(abundance_asv, file.path('intermediate_data', 'abundance_asv.csv'))
write_csv(abundance_otu, file.path('intermediate_data', 'abundance_otu.csv'))
```

## Software used

```{r}
sessioninfo::session_info()
```