### 2021-09-23 05--otu_clustering_threshold.Rmd

---
title: "Clustering ASVs into OTUs"
bibliography: '`r sharedbib::bib_path()`'
output:
  html_document:
    css: style.css
---

```{r setup, include=FALSE}
source('style.R')
```

## Prepare

### Packages used

```{r message=FALSE}
library(dplyr)
library(purrr)
library(furrr)
library(tidyr)
library(readr)
library(ggplot2)
library(sessioninfo)
library(metacoder)
library(vegan)
library(viridis)
library(DT)
library(stringr)
library(qsubmitter)
library(taxize)
library(forcats)
library(parallel)
```

### Parameters

```{r}
opt_thresh_range = seq(0.9, 1, by = 0.001) # The specific thresholds checked during the optimization
opt_min_read_count <- 50 # used in id threshold optimization
opt_cluster_method <- 'size' # 'fast' for length-ordered, 'size' for abundance-ordered
seed <- 1
set.seed(seed)
```

## Read abundance matrix and sample data

```{r}
metadata <- read_csv(file.path('intermediate_data', 'metadata.csv'))
abundance <- read_csv(file.path('intermediate_data', 'abundance_asv.csv'))
print(metadata)
```

Some samples do not have sequences so I will remove those from the metadata table:

```{r}
metadata <- metadata[metadata$sample_id %in% colnames(abundance), ]
```

## Make function to run VSEARCH

I will want to check for an optimal clustering threshold for RPS10 since it is not currently known. 
To do this, I will need to cluster over many thresholds. 
It is important to note that the optimal threshold is relative to the reference sequence database and the clustering algorithm used, therefore this analysis is mostly to get a rough idea of an optimal threshold.

## Find optimal clustering threshold for mock community

First I will filter sample metadata to just mock community and the primers used in the publication.
We are using only the mock community samples since that is the closest thing to a natural community that we know the composition of.
I did something similar with the reference sequences, but they all resulted in too few clusters even if sequences were clustered at 100% (i.e. unique sequences).
On reflection, this is probably because reference databases have more diversity and fewer erroneous sequences than real data from a natural sample would.
Our environmental samples could not be used because we do not know the "correct" number of species.

```{r}
mock_meta <- filter(metadata, dna_type == 'mock2', primer_pair_id %in% c('ITS6/7', 'rps10_Final'))
mock_meta
```

I will need the number of the mock community species added later, so I will import the data for the mock community composition:

```{r}
mc_data <- read_csv(file.path('intermediate_data', 'mock_community.csv'))
```

I will also zero out low abundance ASVs, since that would be done in a normal metabarcoding analysis.
Here we are using a rather high minimum abundance of `r opt_min_read_count` since this is a low-complexity sample, and thus higher read count per species, relative to most natural communities, so erroneous sequence abundance is probably similarly inflated.

```{r}
opt_filtered_abund <- select(abundance, sequence, !!! mock_meta$sample_id)
opt_filtered_abund <- opt_filtered_abund[rowSums(opt_filtered_abund[, mock_meta$sample_id]) > opt_min_read_count, ]
opt_filtered_abund
```

Count the number of OTUs for each sample over a range of clustering thresholds.

```{r}
vserach_cluster <- function(seqs, seq_abund, id_threshold = 0.97, method = "fast") {
  # Check that VSEARCH is installed
  tryCatch(system2("vsearch", args = "--version", stdout = FALSE, stderr = FALSE),
           warning=function(w) {
             stop("vsearch cannot be found on PATH. Is it installed?")
           })

  # Run VSEARCH
  input_fasta_path <- tempfile()
  write_lines(paste0('>', seq_along(seqs), ';size=', seq_abund, '\n', seqs), path = input_fasta_path)
  otu_centroid_path <- tempfile()
  command_args <- paste(paste0("--cluster_", method), 
                        input_fasta_path,
                        "--threads", detectCores() - 1,
                        "--id", id_threshold,
                        "--sizein",
                        "--strand plus",
                        "--fasta_width 0", # 0 = no wrapping in fasta file
                        "--centroids", otu_centroid_path)
  system2("vsearch", args = command_args, stdout = FALSE, stderr = FALSE)

  # Return OTU sequences
  centroids <- read_fasta(otu_centroid_path)
  names(centroids) <- str_match(names(centroids), pattern = 'size=(.+)$')[, 2]
  return(centroids)
}

opt_thresh_data <- tibble(thresh = opt_thresh_range)
opt_thresh_data[mock_meta$locus] <- map(mock_meta$sample_id, function(sample_id) {
  map_dbl(opt_thresh_range, function(threshold) {
    has_seq <- opt_filtered_abund[[sample_id]] > 0
    length(vserach_cluster(seqs = opt_filtered_abund$sequence[has_seq],
                           seq_abund = opt_filtered_abund[[sample_id]][has_seq],
                           id_threshold = threshold, 
                           method = opt_cluster_method))
  })
})
opt_thresh_data
```

find the range of thresholds that return the correct value

```{r}
valid_thresh_range <- do.call(rbind, map(mock_meta$locus, function(l) {
  out <- range(opt_thresh_data$thresh[abs(opt_thresh_data[[l]] - nrow(mc_data)) <= 1])
  names(out) <- c('min_thresh', 'max_thresh')
  return(out)
})) %>% as_tibble()
valid_thresh_range <- as_tibble(valid_thresh_range) %>%
  mutate(locus = mock_meta$locus) %>%
  select(locus, min_thresh, max_thresh)
valid_thresh_range$min_count <- min(opt_thresh_data[mock_meta$locus])
valid_thresh_range$max_count <- max(opt_thresh_data[mock_meta$locus])
valid_thresh_range
```

Now plot

```{r}
opt_thresh_plot <- opt_thresh_data %>%
  gather(key = 'locus', value = 'otu_count', !!! mock_meta$locus) %>%
  mutate(thresh = thresh * 100) %>%
  ggplot() +
  geom_hline(yintercept = nrow(mc_data), linetype = "dashed") +
  geom_rect(aes(xmin = min_thresh * 100, xmax = max_thresh * 100, ymin = min_count, ymax = max_count), data = valid_thresh_range, alpha = 0.5) +
  geom_line(aes(x = thresh, y = otu_count)) +
  scale_x_continuous(breaks = seq(min(opt_thresh_data$thresh), max(opt_thresh_data$thresh), by = 0.01) * 100, 
                     minor_breaks = NULL) +
  scale_y_continuous(minor_breaks = 1:50) +
  facet_wrap(. ~ locus) +
  labs(x = 'PID Clustering Threshold', y = 'OTU Count', fill = '') 
ggsave(opt_thresh_plot, filename = 'optimal_clust_thresh_mock_comm.pdf', path = file.path('results'), width = 10, height = 5)
opt_thresh_plot
```