### 2021-09-23 07--nontarget_amplification.Rmd

---
title: "Non-target amplification"
bibliography: '`r sharedbib::bib_path()`'
output:
  html_document:
    css: style.css
---

```{r setup, include=FALSE}
source('style.R')
```

## Prepare

### Packages used

```{r message=FALSE}
library(dplyr)
library(purrr)
library(furrr)
library(tidyr)
library(readr)
library(ggplot2)
library(sessioninfo)
library(metacoder)
library(vegan)
library(viridis)
library(DT)
library(stringr)
library(qsubmitter)
library(taxize)
library(forcats)
library(ips)
library(insect)
library(ape)
library(phangorn)
```

### Parameters

```{r}
minimum_read_count <- 10
seed <- 1
set.seed(seed)
```

## Detecting non-target amplification

Since the reference databases will not have all the non-targets, it will be hard to detect non-target amplification.
I will use blast against Genbank nt.
Genbank is not curated, but it should be good enough to reliably assign a phylum.

```{r}
abundance_asv <- read_csv(file.path('intermediate_data', 'abundance_asv.csv'))
abundance_asv
abundance_otu <- read_csv(file.path('intermediate_data', 'abundance_otu.csv'))
abundance_otu
metadata <- read_csv(file.path('intermediate_data', 'metadata.csv'))
metadata
```

### Select only environmental samples

Non-target DNA is likely to be found the in the environmental samples, so I will restrict this analysis to those.

```{r}
metadata <- filter(metadata,
                   dna_type %in% c("leaf", "WR.soil", "ag.soil", "drip", "Pan.soil"),
                   primer_pair_id %in% c('rps10_Final', 'ITS6/7'))
metadata
abundance_asv <- select(abundance_asv, sequence:taxonomy, !!! metadata$sample_id)
abundance_asv
abundance_otu <- select(abundance_otu, sequence:taxonomy, !!! metadata$sample_id)
abundance_otu
```

### Filtering out low abundance_asv ASVs

Most low-abundance_asv ASVs would normally be filtered out during most analyses, so I will do that to.

```{r}
abundance_asv <- abundance_asv[rowSums(abundance_asv[, metadata$sample_id]) >= minimum_read_count, ]
abundance_asv
abundance_otu <- abundance_otu[rowSums(abundance_otu[, metadata$sample_id]) >= minimum_read_count, ]
abundance_otu
```

### Group ASVs into target and non-target

```{r}
group_key <- c(Oomycetes = 'Oomycetes|Oomycota', Fungi = 'Fungi', Bacteria = 'Bacteria', Plant = 'Viridiplantae', 
               Archaea = 'Archaea', Animals = 'Metazoa', Virus = 'Viruses', Protist = 'Jakobida|Alveolata|Euglenozoa|Amoebozoa')

groups_target_nontarget <- function(abundance) {
  map_chr(abundance$blast_tax, function(x) {
    if (is.na(x)) {
      return('Unknown')
    }
    is_org <- map_lgl(group_key, grepl, x = x)
    if (sum(is_org) > 1) {
      stop('multiple matches')
    } else if (sum(is_org) == 1) {
      return(names(group_key)[is_org])
    } else {
      return('Other')
    }
  })
}
abundance_asv$group <- groups_target_nontarget(abundance_asv)
abundance_otu$group <- groups_target_nontarget(abundance_otu)
```

### Tally number of reads/ASVs per sample in each group

I will make a table with the numbers of ASVs, reads, and the proportion of reads for each group.
First the ASV counts:

```{r}
groups <- unique(abundance_asv$group)

count_seqs <- function(abundance, measure) {
  counts <- map(metadata$sample_id, function(id) {
    map_dbl(groups, function(group) {
      is_group <- abundance$group == group & abundance[[id]] > 0
      sum(is_group)
    })
  })
  names(counts) <- metadata$sample_id
  counts <- as_tibble(c(group = list(groups), 
                        measure = list(rep(measure, length(group_key))),
                        counts))
  counts
}
asv_counts <- count_seqs(abundance_asv, 'ASVs')
otu_counts <- count_seqs(abundance_otu, 'OTUs')
```

Then read counts

```{r}
count_reads <- function(abundance) {
  read_counts <- map(metadata$sample_id, function(id) {
    map_dbl(groups, function(group) {
      is_group <- abundance$group == group & abundance[[id]] > 0
      sum(abundance[[id]][is_group])
    })
  })
  names(read_counts) <- metadata$sample_id
  read_counts <- as_tibble(c(group = list(groups), 
                             measure = list(rep('read count', length(group_key))),
                             read_counts))
  read_counts
}
read_counts_asv <- count_reads(abundance_asv)
read_counts_otu <- count_reads(abundance_otu)
```

And then I can convert those count to proportions:

```{r}
convert_to_prop <- function(read_counts, abundance) {
  read_props <- read_counts
  read_props[metadata$sample_id] <- map(metadata$sample_id, function(id) {
    read_props[[id]] / sum(abundance[[id]])
  })
  read_props$measure <- 'Reads'
  read_props
}
read_props_asv <- convert_to_prop(read_counts_asv, abundance_asv)
read_props_otu <- convert_to_prop(read_counts_otu, abundance_otu)
```

### Plot barchart of counts

Finally I can combine these into a single table and add some metadata

```{r}
group_abund <- bind_rows(otu_counts, asv_counts, read_counts_asv, read_props_asv) %>%
  gather(key = 'sample_id', value = 'abund', !!! metadata$sample_id) %>%
  left_join(metadata, by = 'sample_id')
group_abund
```

I will calculate the proportion of ASVs and OTUs assigned to each group for each locus

```{r}
group_props <- group_abund %>%
  group_by(measure, locus) %>%
  mutate(sum_abund = sum(abund)) %>%
  group_by(group, .add = TRUE) %>%
  summarise(prop = round(sum(abund) / sum_abund[1], digits = 4)) %>%
  filter(measure %in% c("ASVs", "OTUs", "read count")) 
datatable(group_props)
```

Check that proportions sum to 1 (there will be a bit of rounding error):

```{r}
group_props %>%
  summarise(sum = sum(prop))
```

Lets take a look at what proportion of oomycetes was found for each metric:

```{r}
filter(group_props, group == "Oomycetes")
```

Finally, lets plot this information:

```{r fig.width=5, fig.height=5}
nontarget_plot <- group_abund %>% 
  filter(measure %in% c('Reads', 'ASVs', 'OTUs')) %>%
  mutate(dna_type = fct_collapse(dna_type,
                                 Water = c('drip'),
                                 Plant = c('leaf'),
                                 Soil = c('ag.soil', 'Pan.soil', 'WR.soil')),
         measure = factor(measure, levels = c('Reads', 'ASVs', 'OTUs'), ordered = TRUE),
         locus = ordered(c(rps10 = 'rps10', ITS = 'ITS1')[locus], levels = c('rps10', 'ITS1'))) %>%
  group_by(measure, locus, dna_type) %>%
  mutate(abund = abund / sum(abund)) %>%
  mutate(group = fct_collapse(group,
                              Oomycetes = c('Oomycetes'),
                              Fungi = c('Fungi'),
                              Unknown = c('Unknown'),
                              Other = c('Protist', 'Other', 'Plant', 'Bacteria', 'Animals'))) %>%
  mutate(group = factor(group, levels = c("Unknown", "Other", "Fungi", "Oomycetes"), ordered = TRUE)) %>%
  # ggplot(aes(x = locus, y = abund, fill = group)) +
  ggplot(aes(x = dna_type, y = abund, fill = group)) +
  geom_bar(stat = 'identity') +
  # facet_grid(measure ~ dna_type) +
  facet_grid(measure ~ locus) +
  scale_fill_viridis_d(begin = 0.8, end = 0.2) +
  labs(x = NULL, y = 'Proportion', fill = NULL) +
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = 'bottom') 
ggsave(nontarget_plot, filename = 'non_target_amplification.pdf', path = 'results', width = 5, height = 5)
nontarget_plot
```

Caption:

Figure #: Target vs non-target amplification using oomycete-specific primers for the ITS1 and RPS10 loci. Reads and ASVs are from a variety of environmental samples, grouped into soil, water, and plant tissue samples. ASV sequences were given a coarse taxonomic assignment based on BLAST searches against the NCBI nucleotide sequence database. Those assigned "Unknown" did not have a match with a E-value of at least 0.001. Sequences in "Other" include plant, animal, bacterial, and protist sequences.

## Look at unknown rps10 sequences

```{r}
unknown_asv_data <- abundance_asv %>% 
  gather(key = 'sample_id', value = 'count', !!! metadata$sample_id) %>%
  left_join(metadata, by = 'sample_id') %>%
  filter(locus == 'rps10', group == 'Unknown', count > 0) %>%
  select(sequence, sample_id, count, dna_type, sample_type)
datatable(unknown_asv_data)
```

It seems a lot these are short:

```{r}
hist(nchar(unknown_asv_data$sequence))
```

And somewhat low abundance:

```{r}
hist(unknown_asv_data$count)
```

The really short and low abundance ones are probably errors, so lets remove those:

```{r}
unknown_asv_data <- filter(unknown_asv_data, nchar(sequence) >= 100, count >= 100)
```

I will get a random sample of oomycete reference sequences and see if they cluster together with any in particular or form separate clades

```{r}
rps10_seqs <- read_fasta(file.path('intermediate_data', 'reference_databases', 'rps10_reference_db.fa'))
names(rps10_seqs) <- gsub(names(rps10_seqs), pattern = ';oodb_[0-9]+;$', replacement = '') %>%
  gsub(pattern = '^.+;', replacement = '') %>%
  gsub(pattern = '_', replacement = ' ')
ref_subsample <- rps10_seqs[!duplicated(names(rps10_seqs))]
ref_subsample <- ref_subsample[sample(length(ref_subsample), 10)]
```

And make a multiple sequence alignment with both the reference sequences and unknown sequences

```{r}
unknown_asv_seqs <- setNames(unknown_asv_data$sequence, paste0('ASV ', seq_along(unknown_asv_data$sequence), ' (', unknown_asv_data$count, ')'))
aligned <- c(ref_subsample, unknown_asv_seqs) %>%
  char2dna() %>%
  mafft(method = 'localpair', exec = 'mafft')
```

It turns out that the unknown sequences are very different from eachother and the reference sequences; so much so that I could not calculate a distance matrix to make a tree.

```{r}
image(aligned)
```

I manually BLASTed a few of the sequences on NCBI's website to verify the BLAST on all of the ASVs worked properly and did not find any close matches.
These seem to be some kind of errors or mispriming against unknown organisms.

## Software used

```{r}
sessioninfo::session_info()
```