### 2021-09-23 09--simulated_pcr_results.Rmd

---
title: "Plotting simulated PCR results"
bibliography: '`r sharedbib::bib_path()`'
output:
  html_document:
    css: style.css
---

```{r setup, include=FALSE}
source('style.R')
```

## Prepare

```{r}
library(ape)
library(metacoder)
library(tibble)
```

## Read in results

These results were generated by Felipe Albornoz using Geneious.
They consist of two FASTA files: the reference sequences used for the simulated PCR and the predicted amplicons.

```{r}
ref_seqs <- read.FASTA(file.path("raw_data", "felipes_sim_pcr_results", "all.ref.seqs.rps10.fasta"))
amp_seqs <- read.FASTA(file.path("raw_data", "felipes_sim_pcr_results", "PCR product.fasta"))
```

I will check that all names in `amp_seqs` are in `ref_seqs`

```{r}
if (! all(names(amp_seqs) %in% names(ref_seqs))) {
  stop("All names in `amp_seqs` must be in `ref_seqs`")
}
```

Save location of named seqs in reference

```{r}
amped_indexes <- match(names(amp_seqs), names(ref_seqs))
```

## Clean up headers

The taxonomic information in the FASTA headers need to be cleaned up to be useful for computational analysis and presentation.

```{r}
clean_headers <- function(headers) {
  # remove _F , _sp
  headers <- sub(headers, pattern = "_F$", replacement = "")
  headers <- sub(headers, pattern = "_sp$", replacement = "")

  # Remove existing classifcaitons and only use seq ID
  headers[grepl(headers, pattern = "_Eukaryota;")] <- gsub(headers[grepl(headers, pattern = "_Eukaryota;")],
                                                           pattern = "_Eukaryota;.+$", replacement = "")

  # Remove everything after species name
  matches <- stringr::str_match(headers, "^([a-zA-Z]+_[a-zA-Z]+).*$")[,2]
  headers[! is.na(matches)] <- matches[! is.na(matches)]

  headers <- sub(headers, pattern = "_sp$", replacement = "")
  headers <- sub(headers, pattern = "_x$", replacement = "")

  # Remove underscores
  # headers <- gsub(headers, pattern = "([a-zA-Z]{3,})_", replacement = "\\1 ")

  headers <- sub(headers, pattern = "Peronosppora", replacement = "Peronospora")
  headers <- sub(headers, pattern = "Scleroderma", replacement = "Sclerospora")

  return(headers)
}
cleaned_headers <- clean_headers(names(ref_seqs))
```

Some headers have species names, others have genbank ids. 
I will read in seq ids and names separately and combine classifications.
Since looking up the taxonomy data can be unreliable depending on internet connection, I will cache the results of this in an .rds file.

Seq IDs:

```{r}
seq_id_obj_data_path <- file.path('intermediate_data', 'seq_id_obj.rds')
if (file.exists(seq_id_obj_data_path)) {
  seq_id_obj <- readRDS(seq_id_obj_data_path)
} else {
  seq_id_obj <- lookup_tax_data(cleaned_headers, type = "seq_id")
  saveRDS(seq_id_obj, file = seq_id_obj_data_path)
}
# seq_id_obj <- remove_redundant_names(seq_id_obj)
seq_id_classes <- classifications(seq_id_obj)[names(seq_id_obj$data$query_data)]
```

Seq names:

```{r}
genus_part <- sub(cleaned_headers, pattern = "^([a-zA-Z]+)_.*$",
                  replacement = "\\1")
species_part <- sub(cleaned_headers, pattern = "^[a-zA-Z]+_([a-zA-Z]+)$",
                  replacement = "\\1")
```

I will also save the results of this database search, since it can take a while and it requires manual input when synonyms are encountered, so cannot be included in an Rmd otherwise.

```{r}
taxon_name_obj_data_path <- file.path('intermediate_data', 'taxon_name_obj.rds')
if (file.exists(taxon_name_obj_data_path)) {
  taxon_name_obj <- readRDS(taxon_name_obj_data_path)
} else {
  taxon_name_obj <- lookup_tax_data(genus_part, type = "taxon_name")
  saveRDS(taxon_name_obj, file = taxon_name_obj_data_path)
}
taxon_name_classes <- classifications(taxon_name_obj)[names(taxon_name_obj$data$query_data)]
```

Combine the classifications from the name and ID searches:

```{r}
taxon_name_classes[taxon_name_classes != "unknown taxon"] <- paste0(taxon_name_classes, ";", species_part)[taxon_name_classes != "unknown taxon"]

combined_class <- unname(seq_id_classes)
combined_class[combined_class == "unknown taxon"] <- taxon_name_classes[combined_class == "unknown taxon"]

```

Convert to taxmap format for plotting:

```{r}
obj <- parse_tax_data(combined_class,
                      datasets = list(original_names = names(ref_seqs)),
                      mappings = c("{{index}}" = "{{index}}"))
```

## Filter out ambigous taxa and atrifacts

This will make the tree cleaner and easier to interpret.

```{r}
cleaned_obj <- obj %>%
  filter_taxa(is_leaf, taxon_names %in% c("Glomus", "Phytophthora"), invert = TRUE) %>%
  filter_taxa(grepl(taxon_names, pattern = 'unclassified', ignore.case = TRUE), invert = TRUE) %>%
  filter_taxa(grepl(taxon_names, pattern = 'endosymbiont', ignore.case = TRUE), invert = TRUE) %>%
  filter_ambiguous_taxa() %>%
  filter_taxa(grepl(taxon_names, pattern =  "^[a-zA-Z ]+$"))

cleaned_obj$data$tax_data <- tibble(taxon_id = names(cleaned_obj$data$tax_data), 
                                    input = cleaned_obj$data$tax_data)
```

## Plot what was amplified

I will find which portions of the taxonomy have all of their leaves amplified ...

```{r}
cleaned_obj$mutate_obs("is_amplified",
               unlist(leaves_apply(cleaned_obj,
                                    function(x) length(x) > 0 && all(x %in% names(amp_seqs)),
                                    value = "original_names")))
amplified_leafs <- cleaned_obj$data$original_names[cleaned_obj$data$original_names %in% names(amp_seqs)]
cleaned_obj$data$is_amplified[names(amplified_leafs)] <- TRUE
```

... and plot those taxa in green:

```{r}
set.seed(5)
cleaned_obj %>% 
  filter_taxa(!is_internode) %>%
  remove_redundant_names() %>%
  heat_tree(node_color = ifelse(is_amplified, "green", "grey"),
            node_color_axis_label = "PID to closest other species",
            node_size = n_obs, 
            node_size_range = c(0.005, 0.025),
            node_label_size_range = c(0.012, 0.018),
            # node_label_size_trans = "log10",
            node_size_axis_label = "Number of sequences",
            node_label = Hmisc::capitalize(taxon_names),
            layout = "da", initial_layout = "re",
            background_color = '#FFFFFF',
            # overlap_avoidance = 2,
            output_file = file.path("results", "rps10_simulated_pcr.pdf"))
```