### 2021-09-23 index.Rmd

---
title: "Introduction"
---

This analysis evaluates the effectiveness of the rps10 locus and the primers we designed for use in Oomycete metabarcoding compared to the current ITS1-based method using ITS6 and ITS7.
The criteria for an ideal barcode and assocaited primers for metabarcoding are:

1) The primers can amplify the gene in all of the target group of organisms.
2) The primers do not amplify the gene in any organisms not in the target group.
3) The sequence amplified can differentiate all members of the target group.
4) The reference database contains sequences for all members of the target group.

We have attempted to evaluate each of these criteria to compared methods using RPS10 and ITS1.
This website is the output of the analysis we used to create the results reported in [insert publication].