### 2021-09-23 index.Rmd
fix website

## Analysis code for "A new oomycete metabarcoding method using the rps10 gene"

Authors: Zachary S. L. Foster, Felipe E Albornoz, Valerie J Fieland, Meredith M Larsen, Frank Andrew Jones, Brett M Tyler, Hai D. T. Nguyen, Treena I Burgess, Carolyn Riddell, Hermann Voglmayr, Frank N Martin, Niklaus J Grunwald

bioRxiv 2021.09.22.460084; doi: https://doi.org/10.1101/2021.09.22.460084

This analysis evaluates the effectiveness of the rps10 locus and the primers we designed for use in Oomycete metabarcoding compared to the current ITS1-based method using ITS6 and ITS7.